Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-04-01 W13report.tex
Final edits before submission to LSST repo.

\subsubsection{Pipeline optimization}

@@

Filtering is performed by convolving the image with its PSF, so the
peak of the filtered image of a given source is:
%
\[Ipeak_{filt} = \sum_{(x,y)}I(x,y) * \phi(x,y)\]
%
Hence the equation for measuring its flux reduces to:
%
\[flux = \frac{Ipeak_{filt}}{\sum_{(x',y')}\phi^2(x',y')}\]

If we define

\[w(x,y) \equiv \frac{\phi(x,y)}{\sum_{(x',y')}\phi^2(x',y')}\]

To calculate the variance, we assume a point source model for the intensity

\[I(x,y) = A\phi(x,y) + \epsilon(x,y)\]

\[\langle flux \rangle = \langle \sum_{(x,y)} w(x,y)(A\phi(x,y) +\epsilon(x,y))\rangle\]

Since $\langle \epsilon(x,y) \rangle = 0$ and $\langle \epsilon(x,y) * \epsilon(x',y') \rangle = \sigma^2(x,y)\delta$,

\[\langle flux \rangle = \sum_{(x,y)} w(x,y)A\phi(x,y)\]
and
\[Var(flux) \equiv \langle(flux - \langle flux \rangle)^2\rangle\]
\[Var(flux) = \left\langle{\left(\sum_{(x,y)} w(x,y)\epsilon(x,y)\right)^2}\right\rangle\]
\[Var(flux) = \sum_{(x,y)} w^2(x,y)\sigma^2(x,y)\]
Substituting $w(x,y)$
\[Var(flux) = \frac{\sum_{(x,y)}\phi^2(x,y)\sigma^2(x,y)}{(\sum_{(x',y')} \phi^2(x,y))^2}\]
This implies that the error $\sigma_{filt}$ on the flux at the location of $Ipeak_{filt}$ is
\[\sigma_{filt} = \frac{\sqrt{(\sum_{(x,y)} \phi^2(x,y)\sigma^2(x,y))}}{\sum_{(x',y')} \phi^2(x,y)}\]
%Similarly, the predicted uncertainty of the flux measurement is:
%
%\[\sigma = \frac{\sqrt{\sum_{x,y}\sigma{^2(x,y)}\phi^2(x,y)}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma{^2(x,y)}$ is the is the per--pixel variance of the
%unfiltered image. This reduces to the following:
%{XXX is this correct? - does it agree with Lupton?}
%\[\sigma = \frac{\sqrt{\sigma peak{_{filt}^2}}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma peak{_{filt}^2}$ is the value of the variance plane,
%after filtering, at the peak of the source.

@@

\clearpage
\begin{verbatim}
for prefilter in (True, False):
    for nGauss in (1,3):
        for degGauss in xrange(1,7):
            for order in xrange(1,7):
                runDiffimPipelin(prefilter, nGauss, degGauss, order)
\end{verbatim}

@@

Specifically, we define:
\begin{equation}
n\prime(>\nu) = n(>k\dot\nu)
\label{correction}
\end{equation}  
For our data we scale {\it down} $\nu$ in the model by $k = 0.975, 0.985 $~and~$ 0.99$ for 
v6866601, v6866602, and v6866603 respectively.  The data plotted with these corrections are shown
in the bottom row of Figure\ref{fig-fp_v_thresh}.  These results suggest our variance estimates 
are too {\it high}, and we are therefore detecting too {\it few} false detections.  This does not agree 
entirely with the analysis of the empirical variance, Section~\ref{sec-varmis} and Table~\ref{tab-variance1}, 
which suggest that our variance planes {\it underestimate} the true variance.  However, this makes clear that 
while we are at the level of the theoretical floor in false detection rate, it is imperative to track the 
variance in our data to at the 1\% level or better to optimize both the false detection rate, and the true detection
efficiency. This will be the subject of the Summer 2013 work.

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false detections than positive false detections.  This drove an investigation 
into the symmetry of detection and measurement on positive/negative sources.  To do this we looked at the 
results of the image difference task and compared it to the same runs when the difference image was inverted
(multiplied by -1) before measurement.  The results are as follows:
\begin{itemize}
\item Centroids of negative sources always have integer values.
\item Whether detected on the original difference image or the inverted image, the {\tt DiaSource} lists are the same length in every
case.  That is to say {\tt len(afwTable.SourceCatalog.readFits(\it{path}))} returns the same value in both cases.  We take this to mean
that the same objects trigger detection in both situations although this cannot be directly varified since the objects with failed centroids are
not the same in both lists.  
\item Centroiding on negative sources fails almost all the time: 99.7\% sources with negative flux
have the {\tt badcentroid} flag set.
\item In some instances psfFlux is stored as 0.0 (not {\tt NaN}) without any centroid or photometry flag set in the {\tt Source} object.  This
happens more than we would expect if it was just a chance event.  This indicates an uncaught failure mode.
\item psfFlux can be measured as NaN with valid centroid values and no other flags set.
\item Fitted psfFlux can have the same sign in both the initial and inverted image in 0.1\% of measurements.
\item Some sources have the {\tt EDGE} bit set at one polarity but not the other.  
  This leads to a difference in the reported number of false positives as we only consider 
  those without the {\tt EDGE} flag set.
\end{itemize}

>>>>

%\subsubsection{Pipeline optimization}

@@

Filtering is performed by convolving the image with its PSF, so the
peak of the filtered image of a given source is:
%
\[Ipeak_{filt} = \sum_{(x,y)}I(x,y) * \phi(x,y).\]
%
Hence the equation for measuring its flux reduces to:
%
\[flux = \frac{Ipeak_{filt}}{\sum_{(x',y')}\phi^2(x',y')}.\]

We further define
%
\[w(x,y) \equiv \frac{\phi(x,y)}{\sum_{(x',y')}\phi^2(x',y')}.\]

To calculate the variance, we assume a point source model for the intensity 
%
\[I(x,y) = A\phi(x,y) + \epsilon(x,y)\]
\[\langle flux \rangle = \langle \sum_{(x,y)} w(x,y)(A\phi(x,y) +\epsilon(x,y))\rangle\]
%
with a white noise contribution $\epsilon(x,y)$.  Since $\langle \epsilon(x,y) \rangle = 0$ and $\langle \epsilon(x,y) * \epsilon(x',y') \rangle = \sigma^2(x,y)\delta(x-x', y-y')$:
%
\[\langle flux \rangle = \sum_{(x,y)} w(x,y)A\phi(x,y)\]
and
\[Var(flux) \equiv \langle(flux - \langle flux \rangle)^2\rangle\]
\[Var(flux) = \left\langle{\left(\sum_{(x,y)} w(x,y)\epsilon(x,y)\right)^2}\right\rangle\]
\[Var(flux) = \sum_{(x,y)} w^2(x,y)\sigma^2(x,y).\]
Substituting $w(x,y)$
\[Var(flux) = \frac{\sum_{(x,y)}\phi^2(x,y)\sigma^2(x,y)}{(\sum_{(x',y')} \phi^2(x,y))^2}.\]
This implies that the error $\sigma_{filt}$ on the flux at the location of $Ipeak_{filt}$ is
\[\sigma_{filt} = \frac{\sqrt{(\sum_{(x,y)} \phi^2(x,y)\sigma^2(x,y))}}{\sum_{(x',y')} \phi^2(x,y)}.\]
%Similarly, the predicted uncertainty of the flux measurement is:
%
%\[\sigma = \frac{\sqrt{\sum_{x,y}\sigma{^2(x,y)}\phi^2(x,y)}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma{^2(x,y)}$ is the is the per--pixel variance of the
%unfiltered image. This reduces to the following:
%{XXX is this correct? - does it agree with Lupton?}
%\[\sigma = \frac{\sqrt{\sigma peak{_{filt}^2}}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma peak{_{filt}^2}$ is the value of the variance plane,
%after filtering, at the peak of the source.

@@

%\clearpage
\begin{verbatim}
for prefilter in (True, False):
    for nGauss in (1,3):
        for degGauss in xrange(1,7):
            for spatialKernelOrder in xrange(1,7):
                runDiffimPipeline(prefilter, nGauss, degGauss, spatialKernelOrder)
\end{verbatim}

@@

Specifically, we define scaling factor $k$ to correct the counts at a given threshold:
\begin{equation}
n_{corrected}(>\nu) = n(> k\nu)
\label{correction}
\end{equation}  
For our data we scale {\it down} $\nu$ in the model by $k = 0.975, 0.985 $~and~$ 0.99$ for 
v6866601, v6866602, and v6866603 respectively.  The data plotted with these corrections are shown
in the bottom row of Figure\ref{fig-fp_v_thresh}.  These results suggest our variance estimates 
are too {\it high}, and we are therefore detecting too {\it few} false detections.  This does not agree 
entirely with the analysis of the empirical variance, Section~\ref{sec-varmis} and Table~\ref{tab-variance1}, 
which suggest that our variance planes {\it underestimate} the true variance.  However, this makes clear that 
while we are at the level of the theoretical floor in false detection rate, it is imperative to track the 
variance in our data to at the 1\% level or better to optimize both the false detection rate, and the true detection
efficiency. This will be the subject of the Summer 2013 work.

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false detections than positive false detections.  This drove an investigation 
into the symmetry of detection and measurement on positive/negative sources.  To do this we looked at the 
results of the image difference task and compared it to the same runs when the difference image was inverted
(multiplied by -1) before measurement.  The results are as follows:
\begin{itemize}
\item Centroids of negative sources always have integer values.
\item Whether detected on the original difference image or the inverted image, the {\tt DiaSource} lists are the same length in every
case.  That is to say {\tt len(afwTable.SourceCatalog.readFits(\it{path}))} returns the same value in both cases.  We take this to mean
that the same objects trigger detection in both situations although this cannot be directly verified since the objects with failed centroids are
not the same in both lists.  
\item Centroiding on negative sources fails almost all the time: 99.7\% sources with negative flux
have the {\tt badcentroid} flag set.
\item In some instances psfFlux is stored as 0.0 (not {\tt NaN}) without any centroid or photometry flag set in the {\tt Source} object.  This
happens more than we would expect if it was just a chance event.  This indicates an uncaught failure mode.
\item psfFlux can be measured as NaN with valid centroid values and no other flags set.
\item Fitted psfFlux can have the same sign in both the initial and inverted image in 0.1\% of measurements.
\item Some sources have the {\tt EDGE} bit set at one polarity but not the other.  
  This leads to a difference in the reported number of false positives as we only consider 
  those without the {\tt EDGE} flag set.
\end{itemize}
